---
title: "Viz"
format: html
editor: visual
---

Load libraries.

```{r}
#| include: false
library(tidyverse)
library(gt)
library(gtExtras)
library(ggthemes)
```

Read in the betting odds data (from Bovada).

```{r}
bets = read_csv("odds.csv")
```

A little bit of data prep.

```{r}
bets = bets %>% mutate(Group_Long = str_c("Group ", Group)) %>%
  select(-Group)
```

Creating a table using the "gt" package.

```{r}
bets %>% 
  gt(
    groupname_col = "Group_Long",
    rowname_col = "Team"
  ) %>% 
  fmt_percent(
    columns = c(`1st`, `2nd`, `3rd`,`4th`),
    decimals = 0) %>% 
  data_color(
    columns = c(`1st`, `2nd`),
    colors = scales::col_numeric(
      c("white", "green3"),
        domain = c(0,1)
        )) %>%
  data_color(
    columns = c(`3rd`,`4th`),
    colors = scales::col_numeric(
      c("white", "red"),
        domain = c(0,1)
        ))
  
```

The default table is very long, so let's create two tables and put them side-by-side.

The first table is for Groups A, B, C, D and the second for the remaining groups.

```{r}
tab1 = bets %>% filter(Group_Long %in% c("Group A", "Group B", "Group C", "Group D")) %>% 
  gt(
    groupname_col = "Group_Long",
    rowname_col = "Team"
  ) %>% 
  fmt_percent(
    columns = c(`1st`, `2nd`, `3rd`,`4th`),
    decimals = 0) %>% 
  data_color(
    columns = c(`1st`, `2nd`),
    colors = scales::col_numeric(
      c("white", "green3"),
        domain = c(0,1)
        )) %>%
  data_color(
    columns = c(`3rd`,`4th`),
    colors = scales::col_numeric(
      c("white", "red"),
        domain = c(0,1)
        )) 
```

```{r}
tab2 = bets %>% filter(Group_Long %in% c("Group E", "Group F", "Group G", "Group H")) %>% 
  gt(
    groupname_col = "Group_Long",
    rowname_col = "Team"
  ) %>% 
  fmt_percent(
    columns = c(`1st`, `2nd`, `3rd`,`4th`),
    decimals = 0) %>% 
  cols_align(
    align = c("center"),
    columns = everything()
        ) %>%
  data_color(
    columns = c(`1st`, `2nd`),
    colors = scales::col_numeric(
      c("white", "green3"),
        domain = c(0,1)
        )) %>%
  data_color(
    columns = c(`3rd`,`4th`),
    colors = scales::col_numeric(
      c("white", "red"),
        domain = c(0,1)
        )) 
```

Create a single table from the two tables.

```{r}
listed_tables <- list(tab1, tab2)

gt_two_column_layout(listed_tables) 
```

Save this table as a PNG file.

```{r}
gt_two_column_layout(listed_tables, output = "save",
                     filename = "two_col.png",
                      vwidth = 600, vheight = 1275)
```

If you prefer to make the cell values text transparent, you can use "00" in front of the cell text color hex code.

```{r}
tab1 = bets %>% filter(Group_Long %in% c("Group A", "Group B", "Group C", "Group D")) %>% 
  gt(
    groupname_col = "Group_Long",
    rowname_col = "Team"
  ) %>% 
  fmt_percent(
    columns = c(`1st`, `2nd`, `3rd`,`4th`),
    decimals = 0) %>% 
  data_color(
    columns = c(`1st`, `2nd`),
    colors = scales::col_numeric(
      c("white", "green3"),
        domain = c(0,1)
        )) %>%
  data_color(
    columns = c(`3rd`,`4th`),
    colors = scales::col_numeric(
      c("white", "red"),
        domain = c(0,1)
        )) %>%
  tab_style(
    style = cell_text(color = "#00000000"),
    locations = cells_body(columns = everything()))

tab2 = bets %>% filter(Group_Long %in% c("Group E", "Group F", "Group G", "Group H")) %>% 
  gt(
    groupname_col = "Group_Long",
    rowname_col = "Team"
  ) %>% 
  fmt_percent(
    columns = c(`1st`, `2nd`, `3rd`,`4th`),
    decimals = 0) %>% 
  cols_align(
    align = c("center"),
    columns = everything()
        ) %>%
  data_color(
    columns = c(`1st`, `2nd`),
    colors = scales::col_numeric(
      c("white", "green3"),
        domain = c(0,1)
        )) %>%
  data_color(
    columns = c(`3rd`,`4th`),
    colors = scales::col_numeric(
      c("white", "red"),
        domain = c(0,1)
        )) %>%
  tab_style(
    style = cell_text(color = "#00000000"),
    locations = cells_body(columns = everything())) 

listed_tables <- list(tab1, tab2)

gt_two_column_layout(listed_tables) 

gt_two_column_layout(listed_tables, output = "save",
                     filename = "two_col2.png",
                      vwidth = 600, vheight = 1275)
```

Experiment with creating stacked bar charts.

Need to pivot data to long format first.

```{r}
bets_long = bets %>% pivot_longer(c(`1st`,`2nd`,`3rd`,`4th`),names_to = "Place", values_to = "Prob")
```

```{r}
ggplot(bets_long, aes(x=as_factor(Team),y=Prob,fill=Place)) + 
  geom_col() + geom_hline(yintercept=0.5,linetype="dashed") + 
  facet_wrap(~Group_Long, nrow = 4, scales = "free_x") +
  scale_fill_manual(values=c("green4", "lightgreen", "lightpink", "red3")) + 
  xlab("Team") + ylab("Probability") + theme_few()
```

Save the chart to a PNG file.

```{r}
ggsave("facets.png", width = 8, height = 5)
```
